Rotation and Lithium Confirmation of a 500 Parsec Halo for the Open Cluster NGC\,2516\footnote{This is paper 3 of the Cluster Difference Imaging Photometric Survey.}

### 2021-07-01 cluster_paper/ms.tex
update plotting label bug
[WIP] staging for review
(not done)

\documentclass[12pt,twocolumn,tighten]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

\newcommand{\nautorotdenominator}{1{,}641} % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987} % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumeratormatching}{507} % plot_autorotation.py, LSP=SPDM period
\newcommand{\ncompstardenominator}{820} % plot_compstar_rotation.py
\newcommand{\ncompstarnumerator}{312} % plot_compstar_rotation.py
\newcommand{\ncompfrac}{38.0\%} % plot_compstar_rotation.py, based on plot_autorotation
\newcommand{\nautovscompstardenominator}{705} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{374} % plot_autorotation.py
\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py

@@

As part of this project, we focus here on a single southern open
cluster, and ask: is the cluster halo truly coeval with the core?  We
chose \cn\ for this analysis because it is young (100-200\,Myr) and
close enough ($d\approx400$\,pc) to facilitate photometric
rotation measurements using TESS.  Its halo was reported by
\citet{kounkel_untangling_2019} to span
$\approx20\times10\times350$\,pc.  We want to know: is this halo real
or is it just a collection of interlopers, foreground and background
stars?  To what extent can we use Gaia alone to reliably identify
age-dated needles in the haystack of field stars?  What are the
implications, observationally and theoretically, if we can identify
the dispersed halos of open clusters?

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across
  position and velocity space.} The core (black) was analyzed by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident
  with the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) is a concatenation of studies by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
  used less restrictive membership assignment algorithms described in
  Appendix~\ref{app:clustering}.  The field sample is randomly drawn
  from an $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
  low volume density of the halo stars makes it difficult to visually
  distinguish the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Position and orbit of NGC\,2516 in the Galaxy.}
  Points are halo members with $\pi/\sigma_\pi>20$; the cluster is
  most elongated viewed top-down ({\it left}). 
  The Sun is at $\{X,Y,Z\} = \{ -8122, 0, +20.8\}\,{\rm pc}$.
  Gray lines trace the past (solid) and future (dashed)
  cluster orbit, and blue arrows denote the mean cluster
  velocity after subtracting the local standard of rest: $\{v_{\rm X},
  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms. 
  \label{fig:XYZ}
	}
\end{figure*}

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.700\textwidth]{f4a.pdf}
		}

    \vspace{-0.7cm}
		\subfloat{
			\includegraphics[width=0.700\textwidth]{f4b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors
  across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
  and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
  automated cleaning (see Section~\ref{subsubsec:cluster}).  The Pleiades
  \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
  \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
  Stars in the core and the halo of NGC\,2516 overlap with the
  Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
  members have longer rotation periods than Pleiades members. Rotation
  periods for a comparison sample of field stars are shown in
  Appendix~\ref{app:compstar}.
  \label{fig:rot}
	}
\end{figure*}
\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.700\textwidth]{f5a.pdf}
		}

    \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.700\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Binarity indicators for rotators in \cn.} Sets
  $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom}) are as in
  Figure~\ref{fig:rot}.  Astrometric binaries ($\mathrm{RUWE}>1.2$;
  11\% of stars) appear in red; photometric binaries ($>$0.3\,mag
  above an empirical isochrone; 20\% of stars) are overplotted in
  orange.  
  Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has 281 stars.
  Dividing these stars into ``slow'' and ``fast'' sequences by eye,
  the fraction of stars showing signs of binarity in the slow and fast
  subsamples are 30\% (63/210) and 56\% (40/71) respectively.
  \label{fig:binarity}
	}
\end{figure*}

@@

We then measured the rotation period from the resulting light curve
using the periodogram implementations in \texttt{astrobase}, namely
the traditional Lomb-Scargle, and the Stellingwerf phase-dispersion
minimization algorithm
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light
from the target star and the sky background \citep{Sullivan_2015}.  This
typically resulted in a circular aperture radius of either 1 or 1.5 pixels.
We recorded the top five periodogram peaks from each method, and their
corresponding powers.  Finally, as a check on crowding, we recorded
the number of stars within the aperture with greater
brightness than the target star, and with brightness within 1.25 and
2.5 TESS magnitudes of the target star.

The resulting rotation periods and periodogram powers are reported,
regardless of detection significance, in Table~\ref{tab:maintable}.
To clean these measurements, we designed 
automated cleaning criteria, which yielded sets we denote
$\mathcal{A}$ and $\mathcal{B}$.  For both sets, we considered light
curves for which the peak Lomb-Scargle periodogram period was below 15
days, the normalized Lomb-Scargle power exceeded 0.08, and for which
no companions with brightness exceeding one-tenth of the target star
were in the aperture ({\it i.e.}, excluding visual binaries
with $\Delta T < 1.25$, according to the Gaia DR2 source catalog).
These limits were chosen after inspecting the light curves visually,
to eliminate low-quality rotation periods while retaining high
completeness.  For Set $\mathcal{A}$, this yielded \nautorotnumerator\
light curves, out of \nautorotdenominator\ stars for which rotation
periods might plausibly have been detected.  For Set $\mathcal{B}$, we
additionally required that the Lomb-Scargle and Stellingwerf
phase-dispersion methods yielded identical periods to within 10\%.
This yielded \nautorotnumeratormatching\ light curves.  Boolean flags
for each set are included in Table~\ref{tab:maintable}.
A set of field stars was also analyzed for comparison.
The results are discussed in Appendix~\ref{app:compstar}.

@@

Figure~\ref{fig:rot} shows the resulting rotation
periods, with Set $\mathcal{A}$ shown on top and Set $\mathcal{B}$ on
the bottom.  While Set $\mathcal{A}$ is more complete than Set
$\mathcal{B}$, this completeness comes at the expense of greater
contamination.  For instance, stars above the slow sequence (perhaps
field contaminants) mostly disappear in Set $\mathcal{B}$.  The
``void'' or ``gap'' underneath the slow sequence is also more
pronounced in Set $\mathcal{B}$.  To facilitate a comparison against
the Pleiades and Praesepe, we used the rotation periods and dereddened
colors from Table~5 of \citet{curtis_rup147_2020}, which drew on data
from \citet{rebull_rotation_2016a} and \citet{douglas_k2_2019}
respectively.  The first order conclusion is that the Pleiades and
\cn\ appear gyrochronologically coeval for colors spanning
$0.5<\bpmrpo<1.2$.  The implications of this comparison for the age of
\cn\ are discussed in Section~\ref{disc:absage}.

Figure~\ref{fig:binarity} shows the same data, with the
points colored by indicators of binarity.  To assess astrometric
binarity, we used the renormalized unit weight error (RUWE) from Gaia
EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
1.2 as an astrometric binary (11\% of the overall cluster sample; see
{\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
photometric binarity, we fitted a spline to Figure~\ref{fig:hr},
fixing the nodes by hand.  We then labeled any points brighter than
0.3\,mag above the cluster sequence as photometric binaries.  This
included 20\% of the overall cluster sample.  Many of the resulting
candidate binaries, though not all of them, appear below the slow
rotation sequence.  In Set $\mathcal{B}$, over the color range where
the slow and fast sequence are present ($0.5<\bpmrpo<1.3$), there are
281 stars, of which 103 (37\%) show signs of binarity.  Dividing these
281 stars into ``fast'' and ``slow'' sequences by eye,  the fraction
of stars showing signs of binarity in the slow and fast subsamples are
30\% (63/210) and 56\% (40/71), respectively.  We speculate on
possible explanations in Section~\ref{disc:rotn}.

@@

\begin{figure*}[t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{f6a.pdf}
			\includegraphics[width=0.5\textwidth]{f6b.pdf}
		}

@@

	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf \cn\ members exist hundreds of parsecs from the core, and up
  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
  2D-tangential velocities for halo members.  Stars with detected
  rotation periods in Set~$\mathcal{A}$ are shown in blue;  halo
  members for which rotation periods should have been detectable, but
  were not detected, are shown in orange.  Stars in the core are shown
  as black points.  In
  the {\it top right}, the 
  ratio of detected to expected stars with rotation periods is shown
  versus 3D separation from
  the core and 2D tangential velocity difference.  Bin widths are
  chosen to enforce the same number of stars in the denominator
  ($N=54$) per bin.  The {\it bottom panels} show the innermost and
  outermost 100 stars with detected rotation periods in position and
  velocity space.
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
comparison, we only show stars with $0.5<\bpmrpo<1.2$ for which our
TESS pipeline succeeded in making light curves.  In other words, the
stars in the base sample are those for which we could have plausibly
detected a rotation period.  This color range is preferred because the
the slow sequence becomes less defined for spectral types later than
$\approx$K4V.  The bins in the histograms (top-right) were chosen to
ensure that the same number of stars were in each bin.  Rotation
periods are detected for 66\% of stars at least 25\,pc from the
cluster core.  The average detection rate inside of 25\,pc, 73\%, is
slightly higher.  Within $\approx$3.5\,pc, the period detection rate of
61\% is relatively low, due to crowding and saturation in the TESS
images.

The lower panels of Figure~\ref{fig:physical_x_rotn} give a second
view of how field star contamination affects the outskirts of the
cluster.  Of the outermost 100 stars in position space with detected
rotation periods, $\approx$8 appear as outliers above the slow
sequence, compared to just a few for the innermost 100 stars.  For the
outermost 100 stars in velocity space, $\approx$13 appear as outliers.
After accounting for the stars that were expected to show rotation
periods and did not, this suggests field contaminant rates of
$\approx$50\% for the outermost cluster members in our target list.

@@

The final approach we took for assessing the youth of the
halo population of \cn\ was an analysis of the Li\,\textsc{I}
6708\,\AA\ doublet.  This type of age measurement has been reviewed by
\citet{soderblom_ages_2010}, and its connection to stellar rotation
has been discussed by \citet{bouvier_lithium-rotation_2020}.
For \cn, two spectroscopic datasets seemed important:
Gaia-ESO \citep{gilmore_gaiaeso_2012} and GALAH
\citep{silva_galah_2015}.  At the time of analysis, Gaia-ESO DR4 and
GALAH DR3 were the most relevant
\citep{randich_gaiaeso_2018,buder_galah_2020}.  The target selection
and results from each survey were as follows.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f10.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Comparison against previous analyses of stellar
  rotation in \cn.} 
  The bottom panel is a concatenation of Set $\mathcal{B}$ with the
  results of \citet{Irwin_NGC2516_2007} (I07) and
  \citet{fritzewski_rotation_2020} (F20).
  I07 and F20 analyzed stars in the core of \cn.
  \label{fig:litcomp}
	}
\end{figure*}

From their rotation period analysis of \cn,
\citet{fritzewski_rotation_2020} demonstrated that between 100 and
200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri, M\,35, M\,50, and
the Pleiades) all show similar rotation period distributions.  The FGK
stars show a ``slow sequence'', a population of ``rapid rotators'',
and in some cases stars in the ``void'' between the two populations
({\it e.g.,}, Psc-Eri and \cn\ show more stars in the void than the
Pleiades).  The late M dwarfs with detected rotation periods tend to
rotate rapidly ($<3$\,days), though there may also be late M dwarfs
with slow ($>20$\,days) rotation periods
\citep{stauffer_rotation_2016}.  Although these late M dwarfs are
beyond the magnitude limit of our analysis, they have been explored in
depth by \citet{Irwin_NGC2516_2007} and
\citet{fritzewski_rotation_2020}.  A comparison of our results is
shown in Figure~\ref{fig:litcomp}.  The bottom panel of this plot
includes, in order of precedence: rotation periods from our analysis,
from ``Class 1 and 2'' \citet{fritzewski_rotation_2020} rotators, and
from the rotators in \citet{Irwin_NGC2516_2007}.  The main
contribution of our study beyond these previous analyses is our
sensitivity across both the core and halo of \cn, and a corresponding
expansion in the number of stars that can be analyzed.  In the
following section, we therefore focus on the origins of the features
in the rotation-color diagram, and whether our expanded sample can
provide any new insights.

>>>>

%\documentclass[12pt,twocolumn,tighten]{aastex63}
\documentclass[12pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex63}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate
\usepackage{epigraph}

@@

\newcommand{\nautorotdenominator}{1{,}641} % plot_autorotation.py, w/ runid=NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumerator}{987} % plot_autorotation.py, For nequal<=0, nclose<=1, LSP>0.08. See /doc/20210205_nequal_and_Prot_vs_color_cleanup.txt
\newcommand{\nautorotnumeratormatching}{892} % plot_autorotation.py, Set A, with manual Prot-color cut
\newcommand{\ncompstardenominator}{820} % plot_compstar_rotation.py
\newcommand{\ncompstarnumerator}{312} % plot_compstar_rotation.py
\newcommand{\ncompfrac}{38.0\%} % plot_compstar_rotation.py, based on plot_autorotation
\newcommand{\nautovscompstardenominator}{705} % plot_autorotation.py, w/ runid=compstar_NGC_2516. For nequal<=0, nclose<=1, LSP>0.08. 
\newcommand{\nautovscompstarnumerator}{374} % plot_autorotation.py
\newcommand{\nautofrac}{53.1\%} % plot_autorotation.py

@@

As part of this project, we focus here on a single southern open
cluster, and ask: is the cluster halo truly coeval with the core?  We
chose \cn\ for this analysis because it is young (100-200\,Myr) and
close enough ($d\approx400$\,pc) to facilitate \replaced{photometric
rotation measurements}{measurements of stellar rotation periods} using TESS.  \added{\citet{healy_stellar_2020}
in fact already used TESS rotation periods to study the stellar inclination
distribution of stars in the core of NGC\,2516.}
\replaced{Its halo}{The cluster's halo however} was reported by
\citet{kounkel_untangling_2019} to span
$\approx20\times10\times350$\,pc\added{, which is a larger volume than
has been considered in previous rotation analyses
\citep[][]{Irwin_NGC2516_2007,fritzewski_rotation_2020,healy_stellar_2020}}.
We want to know: is this halo real or is it just a collection of
interlopers, foreground and background stars?  To what extent can we
use Gaia alone to reliably identify age-dated needles in the haystack
of field stars?  What are the implications, observationally and
theoretically, if we can identify the dispersed halos of open
clusters?

\added{To assess whether the stars in the cluster's halo are the same
ages as stars in the cluster's core, we use three different age-dating
techniques: isochrones, gyrochrones, and lithium depletion
\citep[see][for reviews]{soderblom_ages_2010,soderblom_ages_2014}.
Gyrochronology has now been empirically calibrated for FGK stars with
ages spanning $\approx$10\,Myr to $\approx$2.7\,Gyr \citep[{\it
e.g.},][]{rebull_usco_2018,curtis_temporary_2019,curtis_rup147_2020}.
Lithium equivalent widths can be used for relative age-dating of dwarf
stars over similar age ranges, though for a narrower range of spectral
types \citep{sestito_2005}.  One complication when deriving relative
lithium ages of the K dwarfs comes from the correlation between
lithium abundance and stellar rotation rate, which has recently been
reviewed by \citet{bouvier_lithium-rotation_2020}, and is a subject
that we explore in the latter portions of our analysis.  }

@@

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The dense and diffuse components of NGC\,2516 across
  position and velocity space.} The core (black) was analyzed by
  \citet{cantatgaudin_gaia_2018} using Gaia DR2, and is coincident
  with the visual overdensity of stars canonically accepted as ``the
  cluster''.  The halo (blue) is a concatenation of studies by
  \citet{kounkel_untangling_2019} and \citet{meingast_2021}, which
  used less restrictive membership assignment algorithms described in
  Appendix~\ref{app:clustering}.  The field sample is randomly drawn
  from an $(\alpha, \delta, \pi)$ cube centered on the cluster.  The
  low volume density of the halo stars makes it difficult to visually
  distinguish the field and the halo samples.
  \label{fig:gaia6d}
	}
\end{figure*}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Position and orbit of NGC\,2516 in the Galaxy.}
  Points are halo members with $\pi/\sigma_\pi>20$; the cluster is
  most elongated viewed top-down ({\it left}). 
  The Sun is at $\{X,Y,Z\} = \{ -8122, 0, +20.8\}\,{\rm pc}$.
  Gray lines trace the past (solid) and future (dashed)
  cluster orbit, and blue arrows denote the mean cluster
  velocity after subtracting the local standard of rest: $\{v_{\rm X},
  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms. 
  \label{fig:XYZ}
	}
\end{figure*}

@@

\begin{figure*}[tp!]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.700\textwidth]{f4a.pdf}
		}

    \vspace{-0.7cm}
		\subfloat{
			\includegraphics[width=0.700\textwidth]{f4b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors
  across the core and halo of \cn.} Sets $\mathcal{A}$ ({\it top})
  and $\mathcal{B}$ ({\it bottom}) undergo successive stages of
  automated cleaning \replaced{(see Section~\ref{subsubsec:cluster})}{; Set
  $\mathcal{B}$ is the same as Set $\mathcal{A}$, after selecting stars
  below the dashed line in the top panel (see Section~\ref{subsubsec:cluster})}.  The Pleiades
  \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
  \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
  Stars in the core and the halo of NGC\,2516 overlap with the
  Pleiades at $\bpmrpo$$<$1.2.  From $1.2<\bpmrpo<1.7$, some NGC\,2516
  members have longer rotation periods than Pleiades members. Rotation
  periods for a comparison sample of field stars are shown in
  Appendix~\ref{app:compstar}.
  \label{fig:rot}
	}
\end{figure*}

@@

We then measured the rotation period from the resulting light curve using
the \added{Lomb-Scargle} periodogram implementation\deleted{s} in
\texttt{astrobase}\replaced{, namely the traditional Lomb-Scargle, and
the Stellingwerf phase-dispersion minimization algorithm
\citep{lomb_1976,stellingwerf_period_1978,scargle_studies_1982,stellingwerf_period_2011,bhatti_astrobase_2018}}{\citep{lomb_1976,scargle_studies_1982,bhatti_astrobase_2018}}.
We used the CDIPS aperture radius that, based on theoretical
expectations, was expected to give the optimal balance between light from
the target star and the sky background \citep{Sullivan_2015}.  This
typically resulted in a circular aperture radius of either 1 or 1.5
pixels.  We recorded the top five periodogram peaks\deleted{ from each
method}, and their corresponding powers.  Finally, as a check on
crowding, we recorded the number of stars within the aperture with
greater brightness than the target star, and with brightness within 1.25
and 2.5 TESS magnitudes of the target star.

The resulting rotation periods and periodogram powers are reported,
regardless of detection significance, in Table~\ref{tab:maintable}.
To clean these measurements, we designed 
automated cleaning criteria, which yielded sets we denote
$\mathcal{A}$ and $\mathcal{B}$.  For \replaced{both sets}{Set
$\mathcal{A}$}, we considered light
curves for which the peak Lomb-Scargle periodogram period was below 15
days, the normalized Lomb-Scargle power exceeded 0.08, and for which
no companions with brightness exceeding one-tenth of the target star
were in the aperture ({\it i.e.}, excluding visual binaries
with $\Delta T < 1.25$, according to the Gaia DR2 source catalog).
These limits were chosen after inspecting the light curves visually,
to eliminate low-quality rotation periods while retaining high
completeness.  For Set $\mathcal{A}$, this yielded \nautorotnumerator\
light curves, out of \nautorotdenominator\ stars for which rotation
periods might plausibly have been detected.  For Set $\mathcal{B}$, we
additionally required that the Lomb-Scargle \replaced{and Stellingwerf
phase-dispersion methods yielded identical periods to within 10\%}{period
fell below the boundary drawn in Figure~\ref{fig:rot}, which was
constructed the filter using the Pleiades polynomial from
\citet{curtis_rup147_2020}, with hard cuts at the red and blue ends.}.
This yielded \nautorotnumeratormatching\ light curves\added{in Set
$\mathcal{B}$}. 
Boolean flags for each set are included in Table~\ref{tab:maintable}.  A
set of field stars was also analyzed for comparison.  The results are
discussed in Appendix~\ref{app:compstar}.
\added{The justification for the additional cut in Set $\mathcal{B}$ is
that \cn\ is already known to have few if any stars above the slow
sequence \citep{fritzewski_rotation_2020}.  Stars above the slow sequence
are most likely therefore either not cluster members, or they are cluster
members, but with unresolved binary companions contaminating the rotation
period measurement \citep[{\it
e.g.},][Section~5.1]{stauffer_rotation_2016}.  For the purpose of
assessing which are rotationally consistent with being in the cluster,
this is a viable filter; for the purpose of fully exploring whether any
rotational outliers could be bonafide cluster members, Set $\mathcal{A}$
would be the better cut.}

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
    \includegraphics[width=0.700\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Binarity indicators for rotators in \cn.} \replaced{Sets
  $\mathcal{A}$ ({\it top}) and $\mathcal{B}$ ({\it bottom}) are}{Set
  $\mathcal{B}$ is shown} as in
  Figure~\ref{fig:rot}.  Astrometric binaries ($\mathrm{RUWE}>1.2$;
  11\% of stars) appear in red; photometric binaries ($>$0.3\,mag
  above an empirical isochrone; 20\% of stars) are overplotted in
  orange.  
  Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has \replaced{281}{523} stars.
  Dividing these stars into ``slow'' and ``fast'' sequences by eye,
  the fraction of stars showing signs of binarity in the slow and fast
  subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and
  \replaced{56\% (40/71)}{51\% (68/134)}, respectively.
  \label{fig:binarity}
	}
\end{figure*}

Figure~\ref{fig:rot} shows the resulting rotation periods, with Set
$\mathcal{A}$ shown on top and Set $\mathcal{B}$ on the bottom.  While
Set $\mathcal{A}$ is more complete than Set $\mathcal{B}$, this
completeness comes at the expense of greater contamination.  \deleted{For
instance, stars above the slow sequence (perhaps field contaminants)
mostly disappear in Set $\mathcal{B}$.  The ``void'' or ``gap''
underneath the slow sequence is also more pronounced in Set
$\mathcal{B}$.}  To facilitate a comparison against the Pleiades and
Praesepe, we used the rotation periods and dereddened colors from Table~5
of \citet{curtis_rup147_2020}, which drew on data from
\citet{rebull_rotation_2016a} and \citet{douglas_k2_2019} respectively.
The first order conclusion is that the Pleiades and \cn\ appear
gyrochronologically coeval for colors spanning $0.5<\bpmrpo<1.2$.  The
implications of this comparison for the age of \cn\ are discussed in
Section~\ref{disc:absage}.

\explain{Figure~\ref{fig:binarity} has been modified to account for the
revised definition of Set B.}

Figure~\ref{fig:binarity} shows the same data, with the points colored by
indicators of binarity.  To assess astrometric binarity, we used the
renormalized unit weight error (RUWE) from Gaia EDR3.  For plotting
purposes, we labeled anything with RUWE exceeding 1.2 as an astrometric
binary (11\% of the overall cluster sample; see {\it e.g.},
\citealt{belokurov_unresolved_2020}).  To assess photometric binarity, we
fitted a spline to Figure~\ref{fig:hr}, fixing the nodes by hand.  We
then labeled any points brighter than 0.3\,mag above the cluster sequence
as photometric binaries.  This included 20\% of the overall cluster
sample.  Many of the resulting candidate binaries, though not all of
them, appear below the slow rotation sequence.  In Set $\mathcal{B}$,
over the color range where the slow and fast sequence are present
($0.5<\bpmrpo<1.3$), there are \replaced{281}{523} stars, of which
\replaced{103 (37\%)}{174 (33\%)} show signs of binarity.  Dividing these
\replaced{281}{523} stars into ``fast'' and ``slow'' sequences by eye,
the fraction of stars showing signs of binarity in the slow and fast
subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and \replaced{56\%
(40/71)}{51\% (68/134)}, respectively.  We speculate on possible
explanations in Section~\ref{disc:rotn}.

@@

\begin{figure*}[tp]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.48\textwidth]{f6a.pdf}
			\includegraphics[width=0.5\textwidth]{f6b.pdf}
		}

@@

	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf \cn\ members exist hundreds of parsecs from the core, and up
  to $\approx$5\kms\ from the core in tangential velocity.} {\it Top
  left}: Cartesian positions (as in Figure~\ref{fig:XYZ}) and
  2D-tangential velocities for halo members.  Stars with detected
  rotation periods in \replaced{Set~$\mathcal{A}$}{Set~$\mathcal{B}$} are shown in blue;  halo
  members for which rotation periods should have been detectable, but
  were not detected, are shown in orange.  Stars in the core are shown
  as black points.  In
  the {\it top right}, the 
  ratio of detected to expected stars with rotation periods is shown
  versus 3D separation from
  the core and 2D tangential velocity difference.  Bin widths are
  chosen to enforce the same number of stars in the denominator
  ($N=54$) per bin.  The {\it bottom panels} show the innermost and
  outermost 100 stars with detected rotation periods in position and
  velocity space\added{ (from Set~$\mathcal{A})$}.
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

\explain{This paragraph contained a typo: the sentence containg ``at
least 25\,pc from the cluster core'' was meant to refer to the entire
cluster. }

The results are shown in the top two rows of
Figure~\ref{fig:physical_x_rotn}.  In these figures, to ensure a fair
comparison, we only show stars with $0.5<\bpmrpo<1.2$ for which our TESS
pipeline succeeded in making light curves.  In other words, the stars in
the base sample are those for which we could have plausibly detected a
rotation period.  This color range is preferred because the the slow
sequence becomes less defined for spectral types later than $\approx$K4V.
The bins in the histograms (top-right) were chosen to ensure that the
same number of stars were in each bin.  Rotation periods
\added{consistent with a gyrochronological age of 150\,Myr} are detected
for \replaced{66\%}{68\%} of stars \replaced{at least 25\,pc from the
cluster core}{reported to be in the cluster}.  The average detection rate
inside of 25\,pc, \replaced{73\%}{72\%}, is slightly higher\added{ than
outside of 25\,pc, where the detection rate is 63\%}. Within
$\approx$3.5\,pc, the period detection rate of \replaced{61\%}{59\%} is
relatively low, due to crowding and saturation in the TESS images.

The lower panels of Figure~\ref{fig:physical_x_rotn} give a second
view of how field star contamination affects the outskirts of the
cluster.  Of the outermost 100 stars in position space with detected
rotation periods\added{ in Set $\mathcal{A}$}, $\approx$8 appear as
outliers above the slow sequence, compared to just a few for the
innermost 100 stars.  For the outermost 100 stars in velocity space,
$\approx$13 appear as outliers.  After accounting for the stars that were
expected to show rotation periods and did not, this suggests field
contaminant rates of $\approx$50\% for the outermost cluster members in
our \added{initial} target list.

@@

The final approach we took for assessing the youth of the
halo population of \cn\ was an analysis of the Li\,\textsc{I}
6708\,\AA\ doublet.  \deleted{This type of age measurement has been reviewed by
\citet{soderblom_ages_2010}, and its connection to stellar rotation
has been discussed by \citet{bouvier_lithium-rotation_2020}.}
For \cn, two spectroscopic datasets seemed important:
Gaia-ESO \citep{gilmore_gaiaeso_2012} and GALAH
\citep{silva_galah_2015}.  At the time of analysis, Gaia-ESO DR4 and
GALAH DR3 were the most relevant
\citep{randich_gaiaeso_2018,buder_galah_2020}.  The target selection
and results from each survey were as follows.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f10.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Comparison against previous analyses of stellar
  rotation in \cn.} 
  The bottom panel is a concatenation of Set $\mathcal{B}$ with the
  results of \replaced{\citet{Irwin_NGC2516_2007} (I07) and
  \citet{fritzewski_rotation_2020} (F20)}{\citet{Irwin_NGC2516_2007}
  (I07), \citet{fritzewski_rotation_2020} (F20), and
  \citet{healy_stellar_2020} (HM20)}.
  \replaced{I07 and F20}{I07, F20, and HM20} analyzed stars in the
  core of \cn.
  \label{fig:litcomp}
	}
\end{figure*}

\explain{Figure~\ref{fig:litcomp} has been updated to include rotation
periods from \citet{healy_stellar_2020}.}

From their rotation period analysis of \cn,
\citet{fritzewski_rotation_2020} demonstrated that between 100 and
200\,Myr, young clusters (\cn, Blanco\,1, Psc-Eri, M\,35, M\,50, and
the Pleiades) all show similar rotation period distributions.  The FGK
stars show a ``slow sequence'', a population of ``rapid rotators'',
and in some cases stars in the ``void'' between the two populations
({\it e.g.,}, Psc-Eri and \cn\ show more stars in the void than the
Pleiades).  The late M dwarfs with detected rotation periods tend to
rotate rapidly ($<3$\,days), though there may also be late M dwarfs
with slow ($>20$\,days) rotation periods
\citep{stauffer_rotation_2016}.  Although these late M dwarfs are
beyond the magnitude limit of our analysis, they have been explored in
depth by \citet{Irwin_NGC2516_2007} and
\citet{fritzewski_rotation_2020}.  A comparison of our results is
shown in Figure~\ref{fig:litcomp}.  The bottom panel of this plot
includes, in order of precedence: rotation periods from our analysis,
from \replaced{``Class 1 and 2'' \citet{fritzewski_rotation_2020} rotators, and
from the rotators in \citet{Irwin_NGC2516_2007}}{``Class 1 and
2'' \citet{fritzewski_rotation_2020} sources, from
Table 1 of \citet{healy_stellar_2020}, and
from \citet{Irwin_NGC2516_2007}.}.  The main
contribution of our study beyond these previous analyses is our
sensitivity across both the core and halo of \cn, and a corresponding
expansion in the number of stars that can be analyzed.  In the
following section, we therefore focus on the origins of the features
in the rotation-color diagram, and whether our expanded sample can
provide any new insights.
